---
title: "check line numbers for log files"
format: gfm
---

# check line numbers for log files

I fear that some of the ldsc jobs failed, and that is why i now have NA values in my correlation tibble.

I used `wc -l` to create a text file that contains one row per log file with the first column being the 
number of lines in the log file and the second column being the name of the log file.

Let's read that in and see how many log files have different numbers of lines.

```{r}
dat <- readr::read_table(here::here("analysis", "results", "ldsc_genetic_correlations_fmd_wc-outputs.txt"), 
                         col_names = c("n_lines", "file_name")) |>
        dplyr::filter(n_lines < 100) |>
        dplyr::mutate(job_number = 1:12098)
```

I manually examined the log files with different numbers of lines and found that the ones with 67 lines
have two warning lines added to the usual 65 lines of output. 
here, the warnings, in at least some files, say that the 
h2 value for the trait is out of bounds.

Peculiarly, some of the jobs with fewer than 65 lines have output reporting that they failed to 
calculate the h2 for FMD! It's unclear why this would fail in some cases but not in most cases.

I think that I'll delete the log files with fewer than 65 lines and re-run the jobs.

```{r}
dat |>
    dplyr::filter(n_lines < 65) |>
    dplyr::mutate(file_name = here::here("analysis", "data", "ldsc_genetic_correlations_fmd2", file_name)) |>
    dplyr::pull(file_name) |>
    purrr::map(~file.remove(.x))
```





```{r}
```